Constrained Circulation Ensemble

### 2025-11-14 ms/cce.tex
conclusion

\begin{keypoints}
	\item Our CCE can reproduce the 2020 SWUS drought
	\item Soil water initialization influences drought severity (soil moisture, runoff, temperature extremes), but does not exacerbate the meteorological drought, itself 
	\item Warming from climate change may have exacerbated this drought, but to a small extent
	\item 2090 CCE is much drier than 2090 CESM2-LE, indicating potential drought mitigation from coupled circulation change
\end{keypoints}

@@

We ran multiple ensembles of simulations using CESM2, each consisting of 30 members, with varied initial conditions. For each simulation, winds were nudged to the ERA5 reanalysis for the year 2020. 
Nudging occurred four times per day in each grid cell outside the box bounded by 28$^{\circ}$N, 47$^{\circ}$N, 130$^{\circ}$W, 96$^{\circ}$W (Figure \ref{fig:domain}). This serves to constrain the large-scale circulation, imposing meteorological drought conditions for every ensemble member. Inside the bounding box, the atmosphere evolves freely, allowing for local land-atmosphere coupling, which is necessary for studying the effects of initial conditions.  
Sea surface temperatures (SSTs) were imposed from an ERA5 reanalysis product, and all other forcing (e.g. CO$_2$, aerosols) followed the standard conventions for transient simulations as in the CESM2-LE. 
The simulations were initialized in April, using random initial conditions from the CESM2-LE for all the requisite land and atmosphere states (e.g. temperature, soil moisture, humidity).
Each simulation ran for nine months.

@@

\noindent Things to potentially think about:
\begin{itemize}
\item Why is CCE-2090 so dry? (tk plot AMJ convergence histograms for supp, P-E for now, can complicate later)
\end{itemize}

@@

We have higher confidence in our inference with regard to investigating preconditioning effects using the CCE approach.
Drier initializations were associated with lower ET, lower runoff, higher temperatures, and a weaker carbon sink, serving to significantly exacerbate the drought.
However, initialization did not have a strong influence on precipitation, itself. 
Precipitation in our analyses was generally controlled by the circulation pattern, more than by local water availability.
The CCE approach is particularly useful for investigating the influence of antecedent conditions, because we are able to force drought in all simulations, but also sample a wide range of initial conditions

Ensemble boosting is useful and probably quite prudent

>>>>

\begin{keypoints}
	\item Our constrained circulation ensemble can reproduce the 2020 SWUS drought.
	\item Soil water initialization influences drought severity (soil moisture, runoff, temperature extremes), but does not exacerbate the meteorological drought, itself.
	\item Warming from climate change may have exacerbated the monsoon failure, but to a small extent.
	\item Our 2020 constrained ensemble is comparable to droughts seen in a fully coupled large ensemble, and likewise for an 1850 analogue, but is drier than the fully coupled model in 2090.
\end{keypoints}

@@

We ran multiple ensembles of simulations using CESM2, each consisting of 30 members, with varied initial conditions. For each simulation, winds were nudged to the ERA5 reanalysis for the year 2020. 
Nudging occurred four times per day in each grid cell outside the box bounded by 28$^{\circ}$N, 47$^{\circ}$N, 130$^{\circ}$W, 96$^{\circ}$W (Figure \ref{fig:fig1}a). This serves to constrain the large-scale circulation, imposing meteorological drought conditions for every ensemble member. Inside the bounding box, the atmosphere evolves freely, allowing for local land-atmosphere coupling, which is necessary for studying the effects of initial conditions.  
Sea surface temperatures (SSTs) were imposed from an ERA5 reanalysis product, and all other forcing (e.g. CO$_2$, aerosols) followed the standard conventions for transient simulations as in the CESM2-LE. 
The simulations were initialized in April, using random initial conditions from the CESM2-LE for all the requisite land and atmosphere states (e.g. temperature, soil moisture, humidity).
Each simulation ran for nine months.

@@

@@

We have higher confidence in our inference with regard to investigating preconditioning effects using the CCE approach.
Drier initializations were associated with lower ET, lower runoff, higher temperatures, and a weaker carbon sink, serving to significantly exacerbate the drought.
However, initialization did not have a strong influence on precipitation, itself. 
Precipitation in our analyses was generally controlled by the circulation pattern, more than by local water availability.
The CCE approach is particularly useful for investigating the influence of antecedent conditions, because we are able to force the meteorological drought in all simulations, but also sample a wide range of initial conditions.

Because climate risk tends to be concentrated in extreme events, it is prudent to design experiments to study such events beyond their natural occurrence rate in free-running simulations.
Likewise forced changes to the mean in any given variable may not transpose directly to the extreme quantiles. 
For example, we found that for CESM2, climate change tends to increase ET in our study domain on average, but actually results in a decrease in ET during drought years. 
Frameworks to better resolve extreme events include initial condition large ensembles, AMIP ensembles, branched reinitialization ensemble boosting, and the constrained circulation ensemble.
We find the CCE framework to be a useful tool for investigating extreme events, particularly when the simulations can be analyzed relative to an unconstrained initial condition large ensemble.
